Applying Renormalization Group Theory to the Square Well Liquid

### 2014-05-08 papers/thesis-roth/project.tex
need to commit before I can pull Tested on elliot. Took 6:31.78.
minor changes. committing so I can get some work done at home Tested on quipu. Took 0:26.46.
updates to RG stuff. Not a whole lot of writing, but I fixed some code...I hope Tested on elliot. Took 0:19.56.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory and Methods}

\subsection{Renormalization Group Theory}
Renormalization group (RG) theory is an iterative procedure that was
developed to deal with large-scale fluctuations of the order parameter
near the critical point of a system.

@@

\subsection{Coexistence Curve Algorithm}
Coexistence curves are found by modifying the chemical potential $\mu$
in the grand free energy per volume $\Phi/V$. Grand free energy
density is defined as
\begin{align}
  \frac{\Phi}{V} &= \phi \nonumber \\
                 &= f(T,n) - n\mu \nonumber \\
                 &= f(T,n) - n\frac{d}{dn}f(T,\npart)\ .
\end{align}
It is the value $\npart$ that we adjust to find coexistence curves.

The grand free energy density generally has two distinct minima when
plotted versus density. The liquid will be in liquid-vapor equilibrium
when those minima have the same value for $\phi$. There is a local
maximum between those minima, and the value of the density at that
maximum is used for $\npart$. Find this value at a given temperature
then plot $\phi$ vs $n$ at slightly higher temperature. Find the new
value for $\npart$ such that the two minima have the same value for
$\phi$. (It is helpful to note that if $\npart$ is increased,
$\phi(\nliq)$ increases, and $\phi(\nliq)$ decreases if $\npart$ is
decresed.) The program I wrote (npart.py) uses the previous
temperature's $\npart$ as a ``first guess,'' maximizing
$\phi(T,\npart)$. From there, the program adjusts $\npart$
until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
uses this new value of $\npart$ as the ``first guess'' at the next
higher temperature.

@@

When only including $f^{id}$ and $f^{HS}$ terms, SW and SW+RG match up
nicely. Including the $a_1^{SW}$ and $a_2$ terms doesn't work so
well...

@@

\begin{figure}
  \begin{center}
%  \includegraphics[width=\columnwidth]{figs/SW-RG-compare}
  \end{center}
  \caption{Comparison of free energy density for SW and SW+RG at low temperature}
  \label{fig:SW-RG-compare-loT}
\end{figure}

%% \begin{figure}
%%   \begin{center}
%%   \includegraphics[width=\columnwidth]{figs/}
%%   \end{center}  \caption{Comaprison of SW and RG near the critical point}
%%   \label{fig:SW-RG-compare-hiT}
%% \end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory and Methods}\label{sec:theory}

\subsection{Renormalization Group Theory}\label{subsec:RGT}
Renormalization group (RG) theory is an iterative procedure that was
developed to deal with large-scale fluctuations of the order parameter
near the critical point of a system.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}\label{sec:methods}

\subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
Coexistence curves are found by modifying the chemical potential $\mu$
in the grand free energy per volume $\phi(T,n)/V$. Grand free energy
density is defined as
\begin{align}
  \frac{\Phi(T,n)}{V} &= \phi(T,n) \nonumber \\
                 &= f(T,n) - n\mu \nonumber \\
                 &= f(T,n) - n\frac{df(T,n)}{dn}\bigg|_{\npart}\ .
\end{align}
It is the value $\npart$ that we adjust to find coexistence curves.

The grand free energy density generally has two distinct minima when
plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an example.) The liquid will be in liquid-vapor equilibrium
when those minima have the same value for $\phi(T,n)$. There is a local
maximum between those minima, and the value of the density at that
maximum is used for $\npart$. Find this value at a given temperature
then plot $\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new
value for $\npart$ such that the two minima have the same value for
$\phi(T,n)$. (It is helpful to note that if $\npart$ is increased,
$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$ is
decresed.) The program I wrote uses the previous
temperature's $\npart$ as a ``first guess,'' maximizing
$\phi(T,\npart)$. From there, the program adjusts $\npart$
until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
uses this new value of $\npart$ as the ``first guess'' at $\npart$ for the next
higher temperature.

The algorithm is as follows; start at some temperature $T=T_{start}$, with an intial guess for $\npart$.
\begin{enumerate}
  \item Calculate $\nvap$ and $\nliq$ by minimizing $\phi(T_{start},n)$ over an appropriate region
  \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
  \item Calculate \[\delta\mu = \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\nliq - \nvap}\]
  \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + \delta\mu$
  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
  \item Calculate $\phi(T_{start},\npart^*)$
  \item Re-caluclate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some tolerance (I used $tol=10^{-5}$)
  \item Increase the temperature to $T = T + \delta T$ and repeat
\end{enumerate}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
  \caption{Grand free energy per unit volume, SW, low temp}
  \label{fig:SW-phi-lowT}
\end{figure}

@@

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-lowT}
  \end{center}
  \caption{Free energy density for RG at low temperature}
  \label{fig:SW-RG-compare-lowT}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-highT}
  \end{center}
  \caption{Free energy density for RG near critical temp}
  \label{fig:SW-RG-compare-highT}
\end{figure}
